---
title: "Introduction to MHVG2MTS"
author: "Jules"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to MHVG2MTS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r}
library(MHVG2MTS)
```

## Reading Data

The `MHVG2MTS` package provides two functions for reading data: `read_measures` and `read_measure_set`.

### `read_measures`

This function reads a single file of measures.

```{r}
# Example usage:
# Assuming you have a file named "my_class__layer1_my_measure.txt" in the "data" directory
# measures <- read_measures(dir = "data", class_name = "my_class", tmeasure_name = "my_measure", layer_A = "layer1")
```

### `read_measure_set`

This function reads a set of files of measures.

```{r}
# Example usage:
# Assuming you have files named "my_class_1__layer1_my_measure.txt", "my_class_2__layer1_my_measure.txt", etc. in the "data" directory
# measure_set <- read_measure_set(dir = "data", class_name = "my_class", measure_name = "my_measure", i = 1, layer_A = "layer1")
```

## Plotting Data

The `MHVG2MTS` package provides several functions for plotting data.

### `plot_mts`

This function plots a multivariate time series.

```{r}
# Example usage:
# mts_data <- data.frame(Time = 1:10, Series1 = rnorm(10), Series2 = rnorm(10))
# plot_mts(mts_data, cols = c("red", "blue"))
```

### `plot_acf`

This function plots the autocorrelation function of a time series.

```{r}
# Example usage:
# ts_data <- rnorm(100)
# plot_acf(ts_data)
```

### `plot_ccf`

This function plots the cross-correlation function of two time series.

```{r}
# Example usage:
# ts_a <- rnorm(100)
# ts_b <- rnorm(100)
# plot_ccf(ts_a, ts_b)
```

### `plot_dist`

This function plots a distribution.

```{r}
# Example usage:
# dist_data <- data.frame(k = 1:10, value = rnorm(10), Variable = "A")
# plot_dist(dist_data, cols = "red")
```

### `plot_boxplot_dists`

This function plots a boxplot of distributions.

```{r}
# Example usage:
# freq_data <- data.frame(k = rep(1:5, 2), Freq = rnorm(10), Variable = rep(c("A", "B"), each = 5))
# plot_boxplot_dists(freq_data, cols = c("red", "blue"))
```

### `plot_pointplot_dists`

This function plots a point plot of distributions.

```{r}
# Example usage:
# freq_data <- data.frame(k = rep(1:5, 2), Freq = rnorm(10), Variable = rep(c("A", "B"), each = 5))
# plot_pointplot_dists(freq_data, cols = c("red", "blue"))
```

### `plot_ggplot_models`

This function plots boxplots by class models.

```{r}
# Example usage:
# data <- data.frame(feature = rnorm(10), models = rep(c("A", "B"), each = 5))
# plot_ggplot_models(data, models = data$models, level_models = c("A", "B"), feature = "feature", title = "My Plot", cols = c("red", "blue"))
```

### `plot_pca`

This function plots PCA results.

```{r}
# Example usage:
# pca_data <- prcomp(iris[,1:4])
# plot_pca(pca_data, true_classes = iris$Species, ncp = 2, col = c("red", "blue", "green"))
```

### `plot_boxplot_clust`

This function plots boxplots of clustering evaluation measures.

```{r}
# Example usage:
# df_data <- data.frame(k = rep(2:4, each = 5), Freq = rnorm(15))
# plot_boxplot_clust(df_data)
```

### `plot_clusters`

This function plots clusters.

```{r}
# Example usage:
# data <- iris[,1:4]
# true_classes <- iris$Species
# cluster_fit <- kmeans(data, 3)
# plot_clusters(data, true_classes, cluster_fit, col = c("red", "blue", "green"))
```

### `multiplot`

This function plots multiple ggplots in a grid.

```{r}
# Example usage:
# p1 <- ggplot(iris, aes(x = Sepal.Length, y = Sepal.Width)) + geom_point()
# p2 <- ggplot(iris, aes(x = Petal.Length, y = Petal.Width)) + geom_point()
# multiplot(p1, p2, cols = 2)
```

## Sequence Analysis

The `MHVG2MTS` package provides several functions for analyzing sequences.

### `order_level`

This function reorders a data frame by level.

```{r}
# Example usage:
# data <- data.frame(k = factor(c(3, 1, 2)), value = rnorm(3))
# ordered_data <- order_level(data)
```

### `comp_distribution`

This function computes the distribution of a sequence.

```{r}
# Example usage:
# seq_data <- c(1, 2, 2, 3, 3, 3)
# dist <- comp_distribution(seq_data)
```

### `dist_degree`

This function computes the distribution of a given sample sequence.

```{r}
# Example usage:
# seq_data <- list(list(data.frame(V1 = c(1, 2, 2))), list(data.frame(V1 = c(1, 1, 2))))
# model_names <- c("A", "B")
# distributions <- dist_degree(seq_data, model_names, n_inst = 1, col_idx = 1)
```

### `mean_dist_degree`

This function computes the mean of a distribution set.

```{r}
# Example usage:
# dist_data <- list(data.frame(k = 1:2, A = c(0.5, 0.5), B = c(0.2, 0.8)), data.frame(k = 1:2, A = c(0.6, 0.4), B = c(0.3, 0.7)))
# mean_dist <- mean_dist_degree(dist_data, n_inst = 2)
```

### `sd_dist_degree`

This function computes the standard deviation of a distribution set.

```{r}
# Example usage:
# dist_data <- list(data.frame(k = 1:2, A = c(0.5, 0.5), B = c(0.2, 0.8)), data.frame(k = 1:2, A = c(0.6, 0.4), B = c(0.3, 0.7)))
# sd_dist <- sd_dist_degree(dist_data, n_inst = 2)
```

## Simulation

The `MHVG2MTS` package provides functions for simulating time series from different models.

### `simulate_garch`

This function simulates a bivariate GARCH process.

```{r}
# Example usage:
# omega <- c(0.1, 0.2)
# alpha <- matrix(c(0.1, 0.05, 0.05, 0.1), nrow = 2)
# beta <- matrix(c(0.8, 0.1, 0.1, 0.8), nrow = 2)
# varcov <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
# garch_data <- simulate_garch(t = 100, omega, alpha, beta, varcov)
```

### `simulate_var`

This function simulates a bivariate VAR process.

```{r}
# Example usage:
# const <- c(0.5, 0.5)
# phi <- matrix(c(0.5, 0.2, 0.1, 0.6), nrow = 2)
# cov_mat <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
# var_data <- simulate_var(t = 100, const, phi, cov_mat)
```

### `simulate_wn`

This function simulates a bivariate white noise process.

```{r}
# Example usage:
# cov_mat <- matrix(c(1, 0.5, 0.5, 1), nrow = 2)
# wn_data <- simulate_wn(t = 100, cov_mat)
```

## Utility Functions

The `MHVG2MTS` package provides several utility functions for data manipulation and analysis.

### `melt_variables`

This function melts two data frames.

```{r}
# Example usage:
# df1 <- data.frame(k = 1:2, Y_1 = c(0.5, 0.5))
# df2 <- data.frame(k = 1:2, Y_2 = c(0.2, 0.8))
# melted_df <- melt_variables(df1, df2)
```

### `df_dists`

This function creates a list of data frames with sets of degree distributions.

```{r}
# Example usage:
# list_data <- list(data.frame(k = 1:2, A = c(0.5, 0.5)), data.frame(k = 1:2, A = c(0.6, 0.4)))
# model_names <- c("A")
# dists <- df_dists(list_data, model_names, n_inst = 2, k_f = 1, k_l = 2)
```

### `freq_df`

This function joins all frequencies of a data frame of distributions.

```{r}
# Example usage:
# df_data <- data.frame("2" = c(0.5, 0.6), "3" = c(0.5, 0.4))
# freq <- freq_df(df_data)
```

### `draw_tables`

This function generates a table of degree distributions.

```{r}
# Example usage:
# table_data <- data.frame(A = c(0.5, 0.5), B = c(0.2, 0.8))
# draw_tables(table_data)
```

### `comp_pca`

This function computes PCA.

```{r}
# Example usage:
# data <- iris[,1:4]
# pca_results <- comp_pca(data)
```

### `pca_results`

This function generates a module to compute PCA analysis.

```{r}
# Example usage:
# data <- iris
# pca_res <- pca_results(data, col = c("red", "blue", "green"))
```

### `comp_clusters`

This function computes k-means clustering.

```{r}
# Example usage:
# data_redim <- iris[,1:4]
# k <- 3
# true_classes <- iris$Species
# clusters <- comp_clusters(data_redim, k, true_classes)
```

### `clustering_results`

This function generates a module to compute clustering analysis.

```{r}
# Example usage:
# data <- iris
# pca_res <- comp_pca(data[,1:4])
# pca <- pca_res$pca
# clust_res <- clustering_results(pca_res, pca, data, col = c("red", "blue", "green"), title = "Iris")
```

### `clust_det`

This function determines the best number of clusters.

```{r}
# Example usage:
# pca_data <- iris[,1:4]
# data <- iris
# best_k <- clust_det(pca_data, data, title = "Iris")
```
